### 2021-08-31 paper.Rmd
made histograms density plots

When using a tour, a sequence of $d$-dimensional linear projections are obtained from a $p$-dimensional space, where $d \ll p$. Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest. A projection basis, $A_{p \times d}$, is a matrix that characterizes the direction from which the data are projected.  It needs to satisfy an orthonormality condition, which requires that all the columns in $A$ has unit length and are perpendicular to each other. With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$. The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure. (The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration. In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections. (The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D projection which is plotted in the histogram to the right. The histogram exhibits in the top row clustering of the three species, and correspond to the 1D projection with a high PP index value. The histogram in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have no interesting pattern.

@@

\begin{figure*}
   \centering
         \includegraphics[width=\textwidth]{figures/notation-huber1.pdf}
         \includegraphics[width=\textwidth]{figures/notation-huber2.pdf}
\caption{Huber's plot (left plots) of the penguins data (bill length and depth only). The solid line represents how well the projection onto different directions, separates the three species, and the dashed circle is a reference guide. The histogram (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot. These plots are made using the R package PPtreeViz.}
\label{fig:notation-huber}
\end{figure*}

@@

A tour path, that is a sequence of projections, can be generated by geodesic interpolation between a set of target planes (Figure \ref{fig:tour-path}). Different methods for choosing target planes provide different tour paths. The grand tour [@Asimov1985-xr] is generated using randomly selected target. A guided tour [@Cook1995-ae] generates targets as a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$. A local tour (available originally in [@Buja1986-la] rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$. A little tour (available originally in [@Buja1986-la] navigates through all of the coordinate planes, for example, for 2D projections $\{A_{ik} = 1, A_{lj} = 1, \forall i, j, k, l = 1, ..., p; i\neq j\}$.

@@

Figure \ref{fig:notation-target-gen} shows a representation of 1D tour paths of 5D data, drawn on a PCA space. A grand tour path is on the left and a guided tour path is on the right. The grand tour path should be more wide-spread because it is attempting to show all possible projections, if left to run long enough. The guided tour path should be short, as the projection pursuit optimization zeroes on the optimally interesting projection.

>>>>

When using a tour, a sequence of $d$-dimensional linear projections are obtained from a $p$-dimensional space, where $d \ll p$. Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest. A projection basis, $A_{p \times d}$, is a matrix that characterizes the direction from which the data are projected.  It needs to satisfy an orthonormality condition, which requires that all the columns in $A$ has unit length and are perpendicular to each other. With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$. The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure. (The figures produced with a modified function from the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration. In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections. (The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D data projection which is plotted in the density plot to the right. The density plot in the top row exhibits clustering of the three species, and corresponds to a 1D projection with a high PP index value. The density plot in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have a less interesting pattern.

@@

\begin{figure*}
   \centering
         \includegraphics[width=\textwidth]{figures/notation-huber1.pdf}
         \includegraphics[width=\textwidth]{figures/notation-huber2.pdf}
\caption{Huber's plot (left plots) of the penguins data (bill length and depth only). The solid line represents how well the projection onto different directions, separates the three species, and the dashed circle is a reference guide. The density plots (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot. These plots are made using the R package PPtreeViz.}
\label{fig:notation-huber}
\end{figure*}

@@

A tour path, that is a sequence of projections, can be generated by geodesic interpolation between a set of target planes (Figure \ref{fig:tour-path}). Different methods for choosing target planes provide different tour paths. The grand tour [@Asimov1985-xr] is generated using randomly selected targets. A guided tour [@Cook1995-ae] generates targets as a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$. A local tour (available originally in [@Buja1986-la] rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$. A little tour (available originally in [@Buja1986-la] navigates through all of the coordinate planes, for example, for 2D projections $\{A_{ik} = 1, A_{lj} = 1, \forall i, j, k, l = 1, ..., p; i\neq j\}$.

@@

Figure \ref{fig:notation-target-gen} shows a representation of 1D tour paths of 5D data, drawn on a PCA space. Here PCA provides a convenient 2D representation of the 5D.  A grand tour path is on the left and a guided tour path is on the right. The grand tour path should be more wide-spread because it is attempting to show all possible projections, if left to run long enough. The guided tour path should be short, as the projection pursuit optimization zeroes on the optimally interesting projection. (An interesting aside is that although PCA is used here for creating a summary plot, it can also be considered to be a PP method, where the search finds projections where the data is most varied. There are many different PP indexes that have been proposed over the last few decades. @abdi-williams and @joliffe-cadima-pca provide recent overviews of PCA and @jee2009 provides an overview of PP.)
